---
date: <small>`r Sys.Date()`</small>
author: <small>Ieva Kerseviciute</small>
title: Problem at Hand
---

```{r r-setup, include = FALSE}
# snakemake <- readRDS('.features.Rmd.RDS')
snakemake@source('R/rmd_setup.R')
library(foreach)
library(glue)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)

config <- snakemake@config
scriptdir <- snakemake@scriptdir

feature_length <- config$features$length
min_freq <- config$features$min_freq
max_freq <- config$features$max_freq
freq_step <- config$features$freq_step
nperseg <- config$features$nperseg
relative <- config$features$relative

samples <- data.table::fread(snakemake@input$samples) %>%
  .[ SampleID == snakemake@wildcards$sample ] %>%
  .[ , list(SampleID, Run, Age, Sex, DominantHand) ]

prefix <- snakemake@params$prefix
```

```{python python-setup}
import matplotlib.pyplot as plt
import pandas as pd
import numpy as np
import re
from scipy.signal import welch
from scipy.integrate import simps

with open(f'{r.scriptdir}/python/methods.py', 'r') as file:
    exec(file.read())
```

# Extracting features in sample _`r snakemake@wildcards$sample`_

## Summary

Firstly, the events were cropped into epochs of equal length of `r feature_length` seconds. If it was impossible
to divide an event into equal time segments, the event was cropped from both start and end (e.g., an event of
`r round(feature_length * 12 + feature_length / 2, 2)` seconds would be shortened by
`r round(feature_length / 4, 2)` seconds from both start and end in order to make it `r round(feature_length * 12, 2)`
seconds long, a time interval which is divided into exactly 12 segments of `r feature_length` seconds).

The power spectral density (PSD) was estimated within each epoch. Welch's method was used, with the length
of each segment (`nperseg`) of `r nperseg`. The frequency range from `r min_freq` Hz to `r max_freq` Hz was
divided into `r freq_step` Hz width bands (e.g., `r min_freq`-`r min_freq + freq_step` Hz,
`r min_freq + freq_step`-`r min_freq + freq_step * 2` Hz, and so on), and the area under the PSD curve
was calculated for each frequency band using the Simpson's rule.

The extracted features from all events were joined into a single dataset.
The power line frequencies (50-70 Hz) were removed from extracted feature data. For the remainder of the
features, coefficient of correlation was calculated. For the data that is log-normally distributed an estimate
of coefficient of variation is defined as:

$$\hat{cv} = \sqrt{e^{s^{2}_{ln}} - 1}$$

where $s_{ln}$ is the standard deviation of the data after the natural log transformation.

Principal component analysis was performed to search for both event and feature clustering.

```{r}
feature_key <- foreach(file = snakemake@input$feature_key, .combine = rbind) %do% {
  fread(file)
} %>% unique()

feature_key <- feature_key[ Frequency < 50 - freq_step + 1 | Frequency > 70 ]

feature_matrix <- foreach(file = snakemake@input$feature_matrix, .combine = merge) %do% {
  fread(file)
}

feature_matrix <- feature_matrix[ FeatureID %in% feature_key[ , FeatureID ] ]
feature_matrix_rownames <- feature_matrix[ , FeatureID ]

event_key <- foreach(file = snakemake@input$event_key, .combine = rbind) %do% {
  fread(file)
}

stopifnot(all(feature_matrix[ , FeatureID ] %in% feature_key[ , FeatureID ]))
stopifnot(all(colnames(feature_matrix)[ -1 ] %in% event_key[ , EventID ]))
stopifnot(nrow(feature_matrix) == nrow(feature_key))
stopifnot(ncol(feature_matrix) - 1 == nrow(event_key))
```

**Table 1.** Number of events in all runs.

```{r, include = TRUE}
event_key[ , table(Run) ] %>%
        data.table::as.data.table() %>%
        .[ , RunId := gsub(x = Run, pattern = 'run', replacement = '') %>% as.numeric() ] %>%
        .[ order(RunId) ] %>%
        data.table::setnames('N', 'Number of events') %>%
        .[ , Run := paste('Run', RunId) ] %>%
        .[ , list(Run, `Number of events`) ] %>%
        showTable()
```

```{r}
x <- feature_matrix[ , -1 ] %>%
        setDF(rownames = feature_matrix_rownames) %>%
        as.matrix() %>%
        log()

p1 <- data.table(Bandpower = as.numeric(x)) %>%
        ggplot() +
        geom_histogram(
                aes(x = Bandpower, y = after_stat(density)),
                fill = 'white',
                color = 'black',
                binwidth = 0.25,
                linewidth = 0.25
        ) +
        xlab('Absolute band power, log') +
        ylab('Density') +
        theme_bw(base_size = 9) +
        theme(panel.grid = element_blank())

x <- x %>%
        t() %>%
        scale(center = TRUE, scale = FALSE) %>%
        t()

p2 <- x %>%
        apply(1, \(x) { sd(x)^2 %>% exp() %>% -1 %>% sqrt() }) %>%
        data.table(CV = .) %>%
        ggplot() +
        geom_histogram(
                aes(x = CV, y = after_stat(density)),
                fill = 'white',
                color = 'black',
                binwidth = 0.025,
                linewidth = 0.25
        ) +
        ylab('Density') +
        xlab('Coefficient of variation (CV)') +
        theme_bw(base_size = 9) +
        expand_limits(x = 1) +
        theme(panel.grid = element_blank())

# May take some time
event_pca <- prcomp(t(x), rank. = 2)$x %>%
        as.data.table(keep.rownames = 'EventID') %>%
        merge(event_key) %>%
        .[ , EventType := make.names(Event) %>% factor() %>% relevel(ref = 'X0') ] %>%
        .[ ]

# May take some time
feature_pca <- prcomp(x, rank. = 2)$x %>%
        as.data.table(keep.rownames = 'FeatureID') %>%
        merge(feature_key) %>%
        .[ , Frequency := factor(Frequency) ] %>%
        .[ ]

feature_annotation <- feature_pca[ , list(PC1 = median(PC1), PC2 = median(PC2)), by = Frequency ]

p3 <- event_pca %>%
        ggplot() +
        geom_point(aes(x = PC1, y = PC2, color = EventType, fill = EventType), size = 1, alpha = 0.5) +
        theme_bw(base_size = 9) +
        theme(panel.grid = element_blank()) +
        xlab('1st principal component') +
        ylab('2nd principal component') +
        theme(legend.position = 'top') +
        scale_color_discrete(name = 'Event') +
        scale_fill_discrete(name = 'Event')

p4 <- event_pca %>%
        .[ , Run := gsub(x = Run, pattern = 'run', replacement = 'Run ') ] %>%
        ggplot() +
        geom_point(aes(x = PC1, y = PC2, color = Run, fill = Run), size = 1, alpha = 0.5) +
        theme_bw(base_size = 9) +
        theme(panel.grid = element_blank()) +
        xlab('1st principal component') +
        ylab('2nd principal component') +
        theme(legend.position = 'top')

p5 <- feature_pca %>%
        ggplot(aes(x = PC1, y = PC2)) +
        geom_point(aes(fill = Frequency, color = Frequency), size = 1, alpha = 0.75) +
        theme_bw(base_size = 9) +
        theme(panel.grid = element_blank()) +
        xlab('1st principal component') +
        ylab('2nd principal component') +
        theme(legend.position = 'none') +
        geom_label_repel(
                data = feature_annotation,
                aes(x = PC1, y = PC2, label = Frequency),
                min.segment.length = 0,
                inherit.aes = FALSE,
                size = 2.5
        )

```

```{r, include = TRUE, fig.height = 3.6}
ggarrange(
        p1, p2,
        nrow = 1,
        ncol = 2,
        labels = letters[ 1:2 ],
        font.label = list(size = 11)
)
```
```{r, include = TRUE, fig.height = 3.6}
ggarrange(
  p3,
  nrow = 1,
  ncol = 1,
  labels = letters[ 3 ],
  font.label = list(size = 11)
)
```
```{r, include = TRUE, fig.height = 3.6}
ggarrange(
  p4,
  nrow = 1,
  ncol = 1,
  labels = letters[ 4 ],
  font.label = list(size = 11)
)
```
```{r, include = TRUE, fig.height = 3.6}
ggarrange(
        p5,
        nrow = 1,
        ncol = 1,
        labels = letters[ 5 ],
        font.label = list(size = 11)
)
```

**Figure 1.**
**(a)** The distribution of log-transformed feature data.
**(b)** The distribution of coefficient of variation of log-transformed centered feature data.
**(c)** The first two principal components of event PCA. The colors represent different events.
**(d)** The first two principal components of event PCA. The colors represent different runs.
**(e)** The first two principal components of feature PCA. The colors and numbers represent different frequencies.

## Individual runs {.tabset .tabset-pills}

```{r}
pages <- foreach::foreach(run = samples[ , Run ] %>% unique) %do% {
  knitr::knit_child(glue::glue('{snakemake@scriptdir}/reports/components/features.Rmd'), quiet = TRUE, envir = environment())
}
```

```{r, include = TRUE, results = 'asis'}
cat(unlist(pages), sep = '\n')
```
