## Run `r gsub(x = run, pattern = 'run', replacement = '')`

```{python}
prefix = f'{r.prefix}/{r.run}'

raw = read_pickle(f'{prefix}/mne_raw.pkl')
filtered = read_pickle(f'{prefix}/mne_good_channels.pkl')
bad_channels = read_pickle(f'{prefix}/bad_channels.pkl')
```

The raw data were converted into MNE-Python data structures, which were then processed as follows. Firstly, all
channels were scaled using the z-transformation. Then, the data were filtered to remove powerline frequencies
in the ranges 50-70 Hz, 115-125 Hz, 170-190 Hz, and 235-245 Hz using the 40th-order band-stop Butterworth filter.
Finally, the frequencies lower than 4 Hz and higher than 290 Hz were removed using the FIR filter.

```{python, include = TRUE}
raw.compute_psd().plot()
plt.show()
```

**Figure 1.** EEG data before filtering.

```{python, include = TRUE}
filtered.compute_psd().plot()
plt.show()
```

**Figure 2.** EEG data after filtering.

The correlation of power spectral density (PSD) between different channels was calculated. Minimum allowed PSD
correlation was estimated for each run separately by reducing the minimum correlation coefficient until the
number of channels with lower average correlation was less than 25. This algorithm allowed for more robust and
flexible detection of bad quality channels, preventing the removal of good quality data in cases when overall
channel correlation is low (which could be the case with constant correlation cutoff).

```{r, include = TRUE, results = 'asis'}
if (config$filter$dropLowQuality) {
  cat('Low quality channels **were removed** from the data.')
} else {
  cat('Note, that low quality channels **were not removed** from the data.')
}
```

**Table 2.** Channels with low correlation.

```{r, include = TRUE}
data.table::data.table(py$bad_channels) %>% showTable()
```
